---
title: "Analysis of Barrett-Dusy treatment experiments: 2009-2010"
output: author: "Roland Knapp
editor_options: 
  chunk_output_type: console
---
# Analysis of data collected during itraconazole treatment experiments in Dusy and Barrett Lakes basins
## Use *brms* package to conduct regression analyses in a Bayesian framework (default priors used in all analyses)

### Load packages
```{r}
library(EnvStats)
library(bayesplot)
library(brms)
library(loo) 
library(tidyverse)
library(patchwork)
```

### Load data set for analysis of Bd load before, at end, and after itraconazole treatment
```{r}
d <- read_csv(file = here::here("data", "dusybarrett-20092010-captures.csv"))
```

## ANALYSIS 1: Pre-treatment differences in loads between populations assigned as "treated" and "control"
### Create dataset
```{r}
d1 <- d %>%
  filter(date_label == "0_preexperiment") %>% 
  mutate(log10load = log10(bd_load + 1)) %>%
  select(basin, site_id, swab_id, treatment, visit_date, date_label, num_days, stage, bd_load, log10load) %>%
  as_tibble()
```

### Data exploration
```{r}
d1 %>% 
  ggplot(aes(x = date_label, y = log10load, color = stage)) +
    geom_point(position = "jitter") +
    facet_wrap(~ basin + treatment, ncol = 4) +
    stat_summary(fun = median, geom = "point", shape = 18, size = 5, show.legend = FALSE) +
    stat_n_text()

d1 %>%
  ggplot(aes(x = treatment, y = log10load)) +
    geom_boxplot() + 
    facet_wrap(~ basin, ncol = 2) +
    stat_n_text()
```
* Nothing of concern in the data, sample sizes sufficient for analysis.

### Plot distribution of bd_load response variable
```{r}
d1 %>% 
  ggplot(aes(x = bd_load)) +
    geom_histogram()

d1 %>% 
  ggplot(aes(sample = bd_load)) +
    geom_qq() +
    geom_qq_line()
```
* Data are strongly left-skewed, possibly zero-inflated. Model using negative binomial family.

### Fit negative binomial model with 2 population-level effects (treatment category & basin) and interaction
```{r}
d1 <- d1 %>% mutate(bd_load = as.integer(bd_load))
```
```{r}
m1 <- brm(bd_load ~ treatment * basin, 
          data = d1,
          family = negbinomial())
summary(m1)
plot(m1)
plot(conditional_effects(m1), ask = FALSE)
```
* Also fit a zero-inflated model, but it showed little evidence of zero-inflation and was equivalent to negbinomial model based on loo.
* Based on all following code chunks, m1 is best model. 

#### LOO probability integral transform
```{r}
loo_m1 <- loo(m1, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m1)
ppc_loo_pit_overlay(
  y = d1$bd_load, 
  yrep = yrep,
  lw = weights(loo_m1$psis_object))
```
* Some minor deviation from uniformity

### Add group-level effect of site_id
```{r}
m1a <- brm(bd_load ~ treatment * basin +  (1 | site_id), 
           data = d1,
           family = negbinomial())
summary(m1a)
plot((m1a), ask = FALSE)
plot(conditional_effects(m1a), ask = FALSE)
```

### Compare model fit
```{r}
loo(m1, m1a)
```
* No substantial improvement in fit from adding group-level effect (elpd_diff/se_diff < 2). Also seen in LOO-PIT. 

### Add predictor for shape parameter (basin)
```{r}
m1b <- brm(bf(bd_load ~ treatment * basin, shape ~ basin), 
           data = d1,
           family = negbinomial())
summary(m1b)
plot((m1b), ask = FALSE)
plot(conditional_effects(m1b), ask = FALSE)
```
* Effect of shape predictor is unimportant, no improvement in fit indicated by LOO, LOO-PIT.
* m1 is best model. 
* Interpretation: bd_load is similar between populations assigned to treat and control categories and between basins, and relationship between bd_load and treatment category is similar between basins.

## ANALYSIS 2: Immediate effect of treatment on loads (beginning vs. end of treatment)
### Data exploration to facilite dataset creation
```{r}
# number of swabs collected across groups
d %>% select(basin, date_label, num_days, treatment, site_id, stage) %>% 
  filter(date_label == "1_experiment" & treatment == "treated") %>%
  count(basin, treatment, site_id, stage) %>% View()
```
```{r}
# number of swabs collected on each day of treatment
d %>% select(basin, date_label, num_days, treatment, site_id, stage) %>% 
  filter(date_label == "1_experiment" & treatment == "treated") %>%
  count(basin, date_label, treatment, site_id, num_days, stage) %>% 
  arrange(basin, site_id, num_days) %>% View()
```
* Mix of tadpoles and subadults.
* For treated animals, create column to distinguish swabs collected at beginning (num_days: -1, 0, 1) and end of treatment (num_days = 6). 

### Create dataset for analysis
```{r}
d2 <- d %>%
  filter(date_label == "1_experiment" & treatment == "treated") %>%
  mutate(treatcat = if_else(num_days < 6, "begin", "end"),
         across(c("stage", "treatcat", "basin"), as.factor)) %>% 
  select(basin, site_id, swab_id, treatment, treatcat, visit_date, date_label, num_days, stage, bd_load) %>%
  as_tibble()
```

### Data exploration of Bd load patterns across basins, sites, life stages
```{r}
d2 %>% ggplot(aes(x = bd_load)) +
  geom_histogram()
```
```{r}
d2 %>% ggplot(aes(x = treatcat, y = log10(bd_load + 1))) +
  geom_boxplot() +
  facet_grid( ~ basin) +
  stat_n_text()
```
```{r}
d2 %>% ggplot(aes(x = treatcat, y = log10(bd_load + 1))) +
  geom_boxplot() +
  facet_grid( ~ basin + site_id) +
  stat_n_text()
```
```{r}
d2 %>% ggplot(aes(x = treatcat, y = log10(bd_load + 1))) +
  geom_boxplot() +
  facet_grid(stage ~ basin + site_id) +
  stat_n_text()
```
```{r}
# alternative visualization
d2 %>% ggplot(aes(x = treatcat, y = log10(bd_load + 1), color = stage)) +
  geom_point(position = "jitter") +
  facet_grid(~ basin + site_id) +
  stat_n_text()
```
* Bd load is strongly left-skewed and possibly zero-inflated.
* Treatment reduced loads in both basins, but in Dusy load was reduced more than in Barrett. More zeros as a result. 
* Within a basin, sites showed similar patterns (as described above).
* Faceting by life_stage and site_id indicated that many of the subadults swabbed at end of treatment started treatment as tadpoles. As a result, fewer subadults swabbed at beginning of treatment than at end. Only 11469 in Barrett had any subadults at beginning of treatment (and had very few tadpoles). 
* Overall, results suggest model with treatcat * basin term, likely no need to include site_id as random effect. 
* Include stage in model to account for potential stage-specific differences in load.

### Fit negative binomial model with 3 population-level effects plus interaction term
```{r}
d2 <- d2 %>%
  mutate(bd_load = as.integer(bd_load))
```
```{r}
m2 <- brm(bd_load ~ stage + treatcat * basin, 
          data = d2,
          family = negbinomial())
summary(m2)
plot((m2), ask = FALSE)
plot(conditional_effects(m2), ask = FALSE)
```
* Negative effect of treatment on load, no effect of basin or stage, important effect of treat*basin interaction.
* Greater variation in load at beginning than end of treatment, especially for Dusy. 

#### LOO probability integral transform
```{r}
loo2 <- loo(m2, save_psis = TRUE)
yrep <- posterior_predict(m2)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo2$psis_object)
)
```
* Substantial deviation from uniformity. 

### Fit zero-inflated negative binomial model
```{r}
m2a <- brm(bd_load ~ stage + treatcat * basin, 
          data = d2,
          family = zero_inflated_negbinomial())
summary(m2a)
plot((m2a), ask = FALSE)
plot(conditional_effects(m2a), ask = FALSE)
loo(m2, m2a)
```

#### Calculate fit diagnostics
```{r}
loo(m2, m2a)
```
```{r}
loo2a <- loo(m2a, save_psis = TRUE)
yrep <- posterior_predict(m2a)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo2a$psis_object)
)
```
* LOO and LOO-PIT both indicate no improvement in fit (LOO: elpd_diff/se_diff = 1.5).
* However, 95% CI of zi parameter does not include zero, so there is some zi but perhaps not enough to improve model fit. 

### Add group effect (site_id) to negative binomial model
```{r}
m2b <- brm(bd_load ~ stage + treatcat * basin + (1 | site_id), 
           data = d2,
           family = negbinomial(), 
           control = list(adapt_delta = 0.99, max_treedepth = 20), 
           cores = 2)
summary(m2b)
plot((m2b), ask = FALSE)
plot(conditional_effects(m2b), ask = FALSE)
```

#### Compute fit diagnostics
```{r}
loo(m2b, m2)
```
```{r}
loo2b <- loo(m2b, save_psis = TRUE)
yrep <- posterior_predict(m2b)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo2b$psis_object)
)
```
* Adding group effect of site_id had little effect on fit (LOO: elpd_diff/se_diff = 1.1). 

### Add predictors of shape parameter to negative binomial model
```{r}
m2c <- brm(bf(bd_load ~ stage + treatcat * basin, 
              shape ~ stage + treatcat + basin), 
           data = d2,
           family = negbinomial())
summary(m2c)
plot(m2c, ask = FALSE)
plot(conditional_effects(m2c), ask = FALSE)
```
* Adding all three predictors of shape (based on differences in bd_load variation between levels for all three) changed importance of predictors: state, treatcat, and basin all have important effects, as does treatcat*basin interaction.

#### Compute fit diagnostics
```{r}
loo(m2c, m2)
```
```{r}
loo2c <- loo(m2c, save_psis = TRUE)
yrep <- posterior_predict(m2c)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo2c$psis_object)
)
```
* Much better fit than model without predictors of shape parameter (LOO: elpd_diff/se_diff = 5.8). 
* LOO-PIT still shows non-uniformity. 
* Model m2c is best model, but still shows some lack of fit. 

## ANALYSIS 3: Effect of treatment on Bd load in year of treatment and following year
### Create dataset
```{r}
d3 <- d %>%
  filter(num_days > 6) %>% # exclude records from during and before treatment period
  mutate(across(c("stage", "treatment", "basin", "year_std"), as.factor)) %>%
  select(basin, site_id, swab_id, treatment, visit_date, year_std, date_label, num_days, stage, bd_load) %>% as_tibble() 
```

### Data exploration
```{r}
d3 %>% 
  ggplot(aes(x = year_std, y = log10(bd_load + 1), color = stage)) +
  geom_point(position = "jitter") +
  facet_grid(~ basin + treatment + site_id ) +
  stat_n_text(size = 3, angle = 90)
```
```{r}
d3 %>% 
  filter(basin == "dusy") %>%
  ggplot(aes(x = year_std, y = log10(bd_load + 1))) +
  geom_boxplot() +
  facet_grid(~ treatment + site_id + stage) +
  stat_n_text(size = 3, angle = 90)
```
```{r}
d3 %>% 
  filter(basin == "barrett") %>%
  ggplot(aes(x = year_std, y = log10(bd_load + 1))) +
  geom_boxplot() +
  facet_grid(~ treatment + site_id + stage) +
  stat_n_text(size = 3, angle = 90)
```
* General pattern in both basins: (1) subadults have higher loads, (2) loads in control ponds relatively high and constant between years 0 and 1, (3) loads in treated ponds are lower than control ponds in year 0, but increase between years 0 and 1, (4) effects somewhat different between sites, (5) treatment increases probability that bd_load = 0. 
* Suggests the following model: load ~ basin + stage + treatment*year_std + (1 | site_id).

### Distribution of bd_load response variable
```{r}
d3 %>% 
  ggplot(aes(x = bd_load)) +
    geom_histogram()

d3 %>% 
  ggplot(aes(sample = bd_load)) +
    geom_qq() +
    geom_qq_line()
```
* Distribution shows strong negative skew and possible zero-inflation. As with previous datasets, will need to model using negative binomial or zero-inflated negative binomial.

### Fit negative binomial model, population-level predictors only
```{r}
d3 <- d3 %>%
  mutate(bd_load = as.integer(bd_load))
```
```{r}
m3 <- brm(bd_load ~ stage + basin + year_std * treatment,
          data = d3, 
          family = negbinomial())
summary(m3)
plot(m3, ask = FALSE)
plot(conditional_effects(m3), ask = FALSE)
```
* Important predictors of bd_load: stage, treatment, year_std*treatment
* CI differs substantially between stages and treatments.  

#### Compute fit diagnostics
```{r}
loo3 <- loo(m3, save_psis = TRUE)
yrep <- posterior_predict(m3)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3$psis_object))
```
* Some lack of uniformity, but not much. 

### Fit zero-inflated negative binomial model, population-level predictors only
```{r}
m3a <- brm(bd_load ~ stage + basin + year_std * treatment,
          data = d3, 
          family = zero_inflated_negbinomial())
summary(m3a)
plot(m3a, ask = FALSE)
plot(conditional_effects(m3a), ask = FALSE)
```
* Minimal zero-inflation 

#### Compute fit diagnostics
```{r}
loo(m3, m3a)
```
```{r}
loo3a <- loo(m3a, save_psis = TRUE)
yrep <- posterior_predict(m3a)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3a$psis_object))
```
* Despite minimal amount of zero inflation, zero-inflated model provides slightly better fit (loo: elpd_diff/se_diff = 2.5).
* Uniformity slightly improved by zi. 

### Add site_id as group_level predictor
```{r}
m3b <- brm(bd_load ~ stage + basin + year_std * treatment + (1 | site_id),
          data = d3, 
          family = zero_inflated_negbinomial(), 
          cores = 2)
summary(m3b)
plot(m3b, ask = FALSE)
plot(conditional_effects(m3b), ask = FALSE)
```
* Group-level effect of site_id is important. 

#### Compute fit diagnostics
```{r}
loo(m3a, m3b)
```
```{r}
loo3b <- loo(m3b, save_psis = TRUE)
yrep <- posterior_predict(m3b)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3b$psis_object))
```
* Addition of site_id as group-level effect substantially improved fit based on loo (loo: elpd_diff/se_diff = 3.8).
* LOO-PIT shows minimal improvement.

### Add effect of stage and treatment on shape parameter
```{r}
m3c <- brm(bf(bd_load ~ stage + basin + year_std * treatment + (1 | site_id),
              shape ~ stage + basin + year_std + treatment), 
          data = d3, 
          family = zero_inflated_negbinomial(), 
          cores = 2)
summary(m3c)
plot(m3c, ask = FALSE)
plot(conditional_effects(m3c), ask = FALSE)
```
* Stage, basin, year_std, and treatment have important effects on shape parameter. 

#### Compute fit diagnostics
```{r}
loo(m3b, m3c)
```
```{r}
loo3c <- loo(m3c, save_psis = TRUE)
yrep <- posterior_predict(m3c)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3c$psis_object))
```
* Adding predictors for shape substantially improved fit (loo: elpd_diff/se_diff = 11.7). 
* Adding shape predictors reduced deviation from uniformity. 

### Add zero inflation predictors
```{r}
m3d <- brm(bf(bd_load ~ stage + basin + year_std * treatment + (1 | site_id),
              shape ~ stage + basin + year_std + treatment, 
              zi ~ stage + basin + year_std + treatment),
          data = d3, 
          family = zero_inflated_negbinomial(),
          control = list(adapt_delta = 0.99, max_treedepth = 20),  
          cores = 2)
summary(m3d)
plot(m3d, ask = FALSE)
plot(conditional_effects(m3d), ask = FALSE)
```
* Stage, basin, year_std, and treatment have important effects on zero inflation.

#### Compute fit diagnostics
```{r}
loo(m3c, m3d)
```
```{r}
loo3d <- loo(m3d, save_psis = TRUE)
yrep <- posterior_predict(m3d)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3d$psis_object))
```
* Addition of zero-inflation predictors produced little if any improvement in model fit (loo: elpd_diff/se_diff = 2.1). 
* Deviation from uniformity increased relative to m3c. m3c is best model. 

### Create plots of Bd load over 1 year period of experiment for both basins
#### Create labels for plotting
```{r}
d_plot <- d %>% 
  mutate(
    date_label1 = case_when(basin == "barrett" & num_days == "-14" ~ "-14",
                            basin == "barrett" & between(num_days, -1, 5) ~ "-1 to 5",
                            basin == "barrett" & num_days == "6" ~ "6", 
                            basin == "barrett" & between(num_days, 22, 23) ~ "22-23", 
                            basin == "barrett" & between(num_days, 41, 42) ~ "41-42",
                            basin == "barrett" & between(num_days, 365, 366) ~ "365-366",
                            basin == "barrett" & between(num_days, 383, 384) ~ "383-384", 
                            basin == "barrett" & between(num_days, 407, 408) ~ "407-408", 
                            basin == "dusy" & between(num_days, -15, -14) ~ "-15 to -14", 
                            basin == "dusy" & between(num_days, 0, 2) ~ "0-2", 
                            basin == "dusy" & num_days == 6 ~ "6", 
                            basin == "dusy" & between(num_days, 20, 21) ~ "20-21", 
                            basin == "dusy" & between(num_days, 46, 47) ~ "46-47", 
                            basin == "dusy" & between(num_days, 356, 357) ~ "356-357", 
                            basin == "dusy" & between(num_days, 389, 392) ~ "389-392")) %>%
  add_row(basin = "dusy", num_days = 407, date_label1 = "407-408", treatment = "control") %>% # dummy row to match the number of categories in barrett
  add_row(basin = "dusy", num_days = 408, date_label1 = "407-408", treatment = "treated") %>% # dummy row to match the number of categories in barrett
  select(site_id, basin, visit_date, num_days, date_label, date_label1, treatment, stage, bd_load)
```

#### Create Barrett and Dusy plots
```{r}
library(patchwork)
```
```{r}
d_plot1 <- d_plot %>%
  filter(basin == "barrett") %>%
  mutate(date_label1 = fct_relevel(
    date_label1, "-14", "-1 to 5", "6", "22-23", "41-42", "365-366", "383-384", "407-408")) %>%
  ggplot(aes(x = date_label1, y = log10(bd_load + 1), fill = treatment)) +
    geom_boxplot(position = position_dodge(preserve = "single")) + # keep unmatched day 7 same size as others
    labs(title = "(a)",
      x = "Days",
      y = (expression(Bd~load~(log["10"]~copies+1))),
      fill = "Frog group") +
    scale_fill_manual(values=c("forestgreen","deepskyblue2")) + 
    theme_classic() +
    theme(plot.title = element_text(margin = margin(t = 0, b = -30)), 
          axis.title.x = element_blank(), 
          axis.text = element_text(color = "black"),
          legend.position = "top", 
          legend.background = element_rect(size = 0.5, color = "black"))
```
```{r}
d_plot2 <- d_plot %>%
  filter(basin == "dusy") %>%
  mutate(date_label1 = fct_relevel(
    date_label1, "-15 to -14", "0-2", "6", "20-21", "46-47", "356-357", "389-392", "407-408")) %>% 
  ggplot(aes(x = date_label1, y = log10(bd_load + 1), fill = treatment)) +
    geom_boxplot(position = position_dodge(preserve = "single")) + # keep unmatched day 7 same size as others
    labs(title = "(b)",
      x = "Days",
      y = (expression(Bd~load~(log["10"]~copies+1)))) +
    scale_fill_manual(values=c("forestgreen","deepskyblue2")) + 
    theme_classic() +
    theme(axis.text = element_text(color = "black"),
      legend.position = "none")
```
```{r}
plot12 <- d_plot1 / d_plot2
# 2 dummy rows have bd_load = NA, will cause warning message
plot12
```
```{r}
ggsave(here::here("out", "figures", "barrettdusy-20092010-bdload.png"), width = 15, height = 15, units = "cm")
```

## ANALYSIS 4: Effect of treatment on subsequent metamorph counts
### Load data set
```{r}
d4 <- read_csv(file = here::here("data", "dusybarrett-20092010-counts.csv"))
```
```{r}
d4  <- d4 %>% mutate(
  ltadpole = log10(tadpole + 1), 
  across(c("basin", "site_id", "treatment", "year_std"), as.factor))
```

### Data exploration
```{r}
# Distribution of count data
d4 %>% 
  ggplot(aes(x = subadult)) +
    geom_histogram()
```
```{r}
# Subadult counts by year_std 
d4 %>% 
  ggplot(aes(x = year_std, y = log10(subadult + 1))) + 
    geom_boxplot() +
    facet_grid(~ basin + treatment + site_id)
```
```{r}
# Subadult and tadpole counts on every survey data during years 0 and 1
d4 %>% pivot_longer(cols = tadpole:subadult,
                    names_to = "stage",
                    values_to = "count") %>%
ggplot(aes(x = factor(plot_date), y = log10(count + 1), fill = stage)) +
    geom_bar(stat = "identity") + 
    theme(axis.text.x = element_text(angle = 90)) +
  facet_grid(~ basin + treatment + site_id)
```
```{r}
d4 %>% 
  ggplot(aes(x = tadpole, y = subadult, color = treatment)) + 
  geom_point() + 
  facet_grid(~ basin)
```
* Count data are left-skewed and possibly zero-inflated. 
* Subadult counts are somewhat higher in treated ponds than control ponds, but not markedly so. 
* No consistent association between counts of tadpoles versus subadults. 

### Fit poisson model
```{r}
m4 <- brm(subadult ~ basin + ltadpole + year_std * treatment, 
          data = d4,
          family = poisson())
summary(m4)
plot((m4), ask = FALSE)
plot(conditional_effects(m4), ask = FALSE)
```
* Number of tadpoles, year_std, treatment, and year_std*treatment all have important effects in predicted directions.

### Fit negbinomial model
```{r}
m4a <- brm(subadult ~ basin + ltadpole + year_std * treatment, 
          data = d4,
          family = negbinomial(),
          control = list(adapt_delta = 0.99))
summary(m4a)
plot((m4a), ask = FALSE)
plot(conditional_effects(m4a), ask = FALSE)
```

### Compute fit diagnostics
```{r}
loo(m4, m4a)
```
```{r}
loo4a <- loo(m4a, save_psis = TRUE)
yrep <- posterior_predict(m4a)
ppc_loo_pit_overlay(
  y = d4$subadult, 
  yrep = yrep,
  lw = weights(loo4a$psis_object))
```
* Fit of negbinomial model substantially better than poisson (LOO: elpd_diff/se_diff = 3.3). 

### Fit zero-inflated negbinomial model
```{r}
m4b <- brm(subadult ~ basin + ltadpole + year_std * treatment, 
          data = d4,
          family = zero_inflated_negbinomial(),
          control = list(adapt_delta = 0.99))
summary(m4b)
plot((m4b), ask = FALSE)
plot(conditional_effects(m4b), ask = FALSE)
```
* Small amount of zero-inflation, but CI of zi term overlaps zero.

### Compute fit diagnostics
```{r}
loo(m4a, m4b)
```
```{r}
loo4b <- loo(m4a, save_psis = TRUE)
yrep <- posterior_predict(m4b)
ppc_loo_pit_overlay(
  y = d4$subadult, 
  yrep = yrep,
  lw = weights(loo4b$psis_object))
```
* Fit of zero-inflated model is no better than negbinomial (LOO: elpd_diff/se_diff < 1). 

#### Add site_id as group-level effect to model m4a
```{r}
m4c <- brm(bf(subadult ~ basin + ltadpole + year_std * treatment + (1 | site_id)), 
          data = d4,
          family = negbinomial(),
          control = list(adapt_delta = 0.99))
summary(m4c)
plot((m4c), ask = FALSE)
plot(conditional_effects(m4c), ask = FALSE)
```

### Compute fit diagnostics
```{r}
loo(m4a, m4c)
```
```{r}
loo4c <- loo(m4c, save_psis = TRUE)
yrep <- posterior_predict(m4c)
ppc_loo_pit_overlay(
  y = d4$subadult, 
  yrep = yrep,
  lw = weights(loo4c$psis_object))
```
* Slight improvement in fit from adding site_id as group-level effect (LOO: elpd_diff/se_diff = 2.6). 
* Based on all results, m4c is best model. 

#### Add shape predictors to model m4c
```{r}
m4d <- brm(bf(subadult ~ basin + ltadpole + year_std * treatment + (1 | site_id), 
          shape ~ basin + ltadpole + year_std + treatment),
          data = d4,
          family = negbinomial(),
          control = list(adapt_delta = 0.99))
summary(m4d)
plot((m4d), ask = FALSE)
plot(conditional_effects(m4d), ask = FALSE)
```

### Compute fit diagnostics
```{r}
loo(m4c, m4d)
```
```{r}
loo4d <- loo(m4d, save_psis = TRUE)
yrep <- posterior_predict(m4d)
ppc_loo_pit_overlay(
  y = d4$subadult, 
  yrep = yrep,
  lw = weights(loo4d$psis_object))
```
* No improvement of fit from adding shape predictors (LOO: elpd_diff/se_diff < 1)

### Make plots of count data 
```{r}
d4_plot1 <- d4 %>% 
  filter(basin == "barrett") %>% 
  ggplot(aes(x = year_std, y = subadult, fill = treatment)) +
    geom_dotplot(method = "histodot", binaxis = "y", stackdir = "center", stackratio = 1, dotsize = 1.4, stroke = 0.5) + 
    stat_summary(fun = median, geom = "point", shape = 18, size = 2.9, fill = "white", show.legend = FALSE) +
    scale_fill_manual(values = c("forestgreen", "deepskyblue2", "darkgray")) +
    labs(
      title = "(a)",
      x = expression("Year"),
      y = "Number of subadults counted") + 
    facet_grid(~ treatment) + 
    theme_classic() +
    theme(axis.text = element_text(color = "black"),
      legend.position = "none")
```
```{r}
d4_plot2 <- d4 %>% 
  filter(basin == "dusy") %>% 
  ggplot(aes(x = year_std, y = subadult, fill = treatment)) +
    geom_dotplot(method = "histodot", binaxis = "y", stackdir = "center", stackratio = 1, dotsize = 1.4, stroke = 0.5) + 
    stat_summary(fun = median, geom = "point", shape = 18, size = 2.9, fill = "white", show.legend = FALSE) +
    scale_fill_manual(values = c("forestgreen", "deepskyblue2", "darkgray")) +
    labs(
      title = "(b)",
      x = expression("Year"),
      y = "Number of subadults counted") + 
    facet_grid(~ treatment) + 
    theme_classic() +
    theme(axis.text = element_text(color = "black"),
      legend.position = "none")
```
```{r}
d4_plot1 / d4_plot2
```
```{r}
ggsave(here::here("out", "figures", "barrettdusy-yrs01-counts.png"), width = 8, height = 15, units = "cm")
```

