---
title: "dusy-2012-analyses"
author: "Roland Knapp"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Analyses of 2012 Jliv treatment experiment with subadult frogs
## Use *brms* package to conduct regression analyses in a Bayesian framework (default priors used in all analyses)

### Load packages
```{r}
library(EnvStats)
library(brms)
library(loo)
library(bayesplot)
library(lubridate)
library(tidyverse)
```

### Load data
```{r}
d <- read_csv(file = here::here("data", "dusy-2012-swabs.csv"))
```

### Create dataframe for use in datasets developed below
```{r}
diff1 <- as.numeric(as.Date("2012-07-12") - as.Date("2012-07-19"))
diff2 <- as.numeric(as.Date("2012-07-19") - as.Date("2012-07-19"))
diff3 <- as.numeric(as.Date("2012-07-31") - as.Date("2012-07-19"))
diff4 <- as.numeric(as.Date("2012-08-25") - as.Date("2012-07-19"))
diff5 <- as.numeric(as.Date("2012-09-13") - as.Date("2012-07-19"))

days <- c(diff1, diff2, diff3, diff4, diff5)
group_date <- c("2012-07-12", "2012-07-19", "2012-07-31", "2012-08-25", "2012-09-13")

jliv_days <- tibble(group_date, days) %>% 
  mutate(group_date = as.Date(group_date))

d <- d %>% 
  left_join(jliv_days, by = "group_date")
```

### Exploration of full dataset 
#### Distribution of bd load data
```{r}
d %>% 
  ggplot(aes(x = bd_load)) +
    geom_histogram()

d %>% 
  ggplot(aes(sample = bd_load)) + 
    geom_qq() +
    geom_qq_line()

```
* Bd load data are highly left-skewed and appear zero-inflated. 

#### General patterns of Bd load before and after itraconazole/Jliv treatment during summer 2012
```{r}
d %>% 
  mutate(expt_trt = replace(expt_trt, is.na(expt_trt) & expt_or_wild == "wild", "wild")) %>%
  ggplot(aes(x = expt_trt, y = log10(bd_load + 1), fill = expt_trt)) +
    geom_dotplot(method = "histodot", binaxis = "y", stackdir = "center", stackratio = 0.3, dotsize = 0.5, stroke = 0.5) +
    scale_fill_manual(values = c("forestgreen", "deepskyblue2", "darkgray")) +
    facet_grid(~ group_date) +
    stat_n_text(size = 3) + 
    stat_summary(fun = median, geom = "point", shape = 4, size = 3, show.legend = FALSE) +
    labs(
      x = "Group",
      y = "Bd load (log10(copies + 1)") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90))
```
* Bd loads of subadults assigned to control and itraconazole-treated groups were similar immediately before treatment began
* Itraconazole treatment greatly reduced loads.
* Following Jliv exposure and release of subadults into pond, Bd loads of treated animals increased and by September were similar to loads of wild subadults.
* Control subadults disappeared quickly, resulting in few captures of animals in this group on later dates. 

### ANALYSIS #1: Were Bd loads similar between treated/control groups before itraconazole treatment?

#### Create data set
```{r}
d1  <- d %>% filter(group_date == "2012-07-12") %>%
             mutate(expt_trt = as.factor(expt_trt),
                    days = as.factor(days),
                    bd_load = as.integer(bd_load))  
```
```{r}
# Check distribution of Bd_load
d1 %>% 
  ggplot(aes(x = bd_load)) +
    geom_histogram()
```
* Negative skew, not nearly as strong as with full dataset. No obvious zero-inflation. Negative binomial is reasonable distribution.

#### Fit negative binomial model
```{r}
m1 <- brm(bd_load ~ expt_trt, 
          data = d1,
          family = negbinomial())

summary(m1)
plot(m1)
conditional_effects(m1)
```
* Before treatment, no difference in Bd load between frogs assigned to treated vs control categories.

#### Fit zero inflated negative binomial model
```{r}
m1a <- brm(bd_load ~ expt_trt, 
          data = d1,
          family = zero_inflated_negbinomial())
summary(m1a)
plot(m1a)
conditional_effects(m1a)
```
* No indication of zero-inflation, based on overlap of 95% CI with 0. Use m1. 

#### LOO probability integral transform
```{r}
loo_m1 <- loo(m1, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m1)
ppc_loo_pit_overlay(
  y = d1$bd_load, 
  yrep = yrep,
  lw = weights(loo_m1$psis_object))
```
* Relatively uniform distribution, indicating adequate fit. 

### ANALYSIS #2: Effect of itraconazole treatment on loads (begin vs end of treatment)
#### Create dataset
```{r}
d2 <- d %>% filter(expt_trt == "treated" & (group_date == "2012-07-12" | group_date == "2012-07-19")) %>% 
  mutate(expt_trt = as.factor(expt_trt),
         days = as.factor(days),
         bd_load = as.integer(bd_load))  
```
```{r}
d2 %>% 
  ggplot(aes(x = bd_load)) +
    geom_histogram()
```
* Strong negative skew and zero inflation

#### Fit negative binomial model
```{r}
m2 <- brm(bd_load ~ days,
          data = d2, 
          family = negbinomial())
summary(m2)
plot(m2)
conditional_effects(m2)
```
* Strong negative effect of treatment on Bd load, much less variation after treatment. 

#### Fit a zero-inflated negative binomial model
```{r}
m2a <- brm(bd_load ~ days,
          data = d2, 
          family = zero_inflated_negbinomial())
summary(m2a)
plot(m2a)
conditional_effects(m2a)
```
* Results indicate substantial zero-inflation.

#### Compare fit of model with and without zero inflation
```{r}
loo(m2, m2a)
```
* Zero-inflated model (m2a) is substantially better (elpd_diff/se_diff = 3.8)

#### LOO probability integral transform
```{r}
loo_m2a <- loo(m2a, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m2a)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo_m2a$psis_object))
```
* Relatively uniform distribution except at lowest x-axis values.

#### Add group_date as a predictor of zero inflation
```{r}
m2b <- brm(bf(bd_load ~ days, zi ~ days),
           data = d2,
           family = zero_inflated_negbinomial())
summary(m2b)
plot(m2b)
conditional_effects(m2b)
```
* Itraconazole treatment increased the probability of bd_load = 0

#### Compare fit of models with and without zero inflation predictor
```{r}
loo(m2a, m2b)
```
* Fit of model with zi predictor is somewhat better than model without, but not substantially so (elpd_diff/se_diff = 2.5). 

#### LOO probability integral transform
```{r}
loo_m2b <- loo(m2b, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m2b)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo_m2b$psis_object))
```
* PIT figure showed no improvement with addition of zi predictor. 
* Addition of shape predictor to m2a indicated that shape predictor was relatively unimportant. Model m2a is best fit. 

### ANALYSIS #3: Temporal patterns of Jliv concentration on treated, control, and wild subadults
#### General patterns of Jliv load
```{r}
# Distribution of Jliv concentrations in 2012, excluding records from prior to Jliv exposure
d %>%
  filter(group_date > "2012-07-19") %>%
  ggplot(aes(x = jliv_ge)) +
        geom_histogram()
```
* Highly left-skewed and possible zero-inflation. Negative binomial distribution is reasonable. 

#### Exploration of Jliv concentrations through time
```{r}
# Jliv concentration on control, treated, and wild subadults, faceted by date
d %>%
  mutate(expt_trt = replace(expt_trt, is.na(expt_trt) & expt_or_wild == "wild", "wild")) %>%
  filter(group_date > "2012-07-12") %>%
  ggplot(aes(x = expt_trt, y = log10(jliv_ge + 1), fill = expt_trt)) +
    geom_boxplot() +
    scale_fill_manual(values = c("forestgreen", "deepskyblue2", "darkgray")) +
    facet_grid(~ group_date) +
    stat_n_text() +
    labs(
      x = "Group",
      y = expression("Concentration of"~italic("J. lividum")~"(genomic equivalents)")) +
    theme(axis.text.x = element_text(angle = 90))
```
* Very low Jliv concentration on "treated" group immediately prior to Jliv exposure. 
* 2 weeks following Jliv exposure of "treated" group, treated, control, and wild groups all had substantial jliv concentrations.
* Decreased at similar rate across groups, returned to pre-exposure concentrations by mid-September

#### Create dataset
```{r}
d3 <- d %>%
  mutate(
    frog_group = expt_trt,
    frog_group = replace(frog_group, is.na(frog_group) & expt_or_wild == "wild", "wild")) %>% 
  filter(group_date > "2012-07-19") %>%
  mutate(frog_group = as.factor(frog_group), 
         jliv_ge = as.integer(jliv_ge))
```

#### Fit negative binomial model
```{r}
m3 <- brm(jliv_ge ~ days,
          data = d3,
          family = negbinomial())
summary(m3)
plot(m3)
conditional_effects(m3)
```
* frog_group not included as predictor in model because there is only 1 sample for control group on 2nd date and none on 3rd date 
* Important effect of days on jliv_ge, with exponential decline of jliv_ge over time 

#### LOO probability integral transform
```{r}
loo_m3 <- loo(m3, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m3)
ppc_loo_pit_overlay(
  y = d3$jliv_ge, 
  yrep = yrep,
  lw = weights(loo_m3$psis_object))
```
* Some lack of fit indicated by departure from uniform distribution at small and large x-values

#### Fit zero inflated negative binomial model
```{r}
m3a <- brm(jliv_ge ~ days,
          data = d3,
          family = zero_inflated_negbinomial())
summary(m3a)
plot(m3a)
conditional_effects(m3a)
```
* No evidence for importance of zero inflation (95% CI includes 0)

#### Compare fit of negbinomial and zi_negbinomial models
```{r}
loo(m3, m3a)
```
* No evidence of zero inflation (elpd_diff/se_diff < 1), use m3. 

#### Fit model with days and treatment group as predictors, exclude control treatment because of small sample sizes
```{r}
d3a <- d3 %>% filter(frog_group != "control")
```
```{r}
m3b <- brm(jliv_ge ~ days * frog_group,
          data = d3a,
          family = negbinomial())
summary(m3b)
plot(m3b)
plot(conditional_effects(m3b), ask = FALSE)
```
* No effect of treated vs wild on Jliv concentration
* No interaction between treated/wild and days 

#### Plot of Jliv concentrations over time 
```{r}
anno_text1 <- tibble(expt_trt = "control", jliv_ge = -0.57, lab = "text",
                    days = factor(56, levels = c("0", "12", "37", "56")))
```
```{r}
facet_names1 <- c('0' = "Day 0",
                 '12' = "Day 12",
                 '37' = "Day 37",
                 '56' = "Day 56")
```
```{r}
d %>%
  mutate(
    expt_trt = replace(expt_trt, is.na(expt_trt) & expt_or_wild == "wild", "wild")) %>%  
  filter(group_date > "2012-07-12") %>% 
  ggplot(aes(x = expt_trt, y = log10(jliv_ge + 1), fill = expt_trt)) + 
    geom_boxplot(width = 0.9) +  
    scale_fill_manual(values = c("forestgreen", "deepskyblue2", "darkgray")) +
    labs(fill = "Frog group", 
         x = "Frog group",
         y = expression(~italic("J. lividum")~"concentration"~(log[10]~("GE + 1")))) + 
    stat_n_text(size = 3) +
    geom_text(data = anno_text1, label = "n=0", size = 3) +
    facet_grid(~ days, labeller = as_labeller(facet_names1)) +
    theme_classic() +
    theme(axis.text = element_text(color = "black", angle = 90),
          axis.title.x = element_text(vjust = 0),
          legend.position = "none")
```
```{r}
ggsave(here::here("out", "figures", "dusy-2012-jliv.png"), width = 15, height = 10, units = "cm")
```

### ANALYSIS #4: Differences in Bd load between itraconazole+Jliv treated and wild subadults 
### Distribution of Bd load data
```{r}
d3a %>%
  ggplot(aes(x = bd_load)) +
    geom_histogram()
```
* Strongly left-skewed, possible zero-inflation.

### Fit negative binomial model
```{r}
m4 <- brm(bd_load ~ days * frog_group,
          data = d3a,
          family = negbinomial())
summary(m4)
plot(m4)
plot(conditional_effects(m4), ask = FALSE)
```
* Important effect on bd_load of days (+), expt_trt (higher for wild), and interaction

#### LOO probability integral transform
```{r}
loo_m4 <- loo(m4, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m4)
ppc_loo_pit_overlay(
  y = d3a$bd_load, 
  yrep = yrep,
  lw = weights(loo_m4$psis_object))
```
* Some lack of fit indicated by deviations from uniform distribution.

#### Fit zero inflated negbinomial model
```{r}
m4a <- brm(bd_load ~ days * frog_group,
          data = d3a,
          family = zero_inflated_negbinomial())
summary(m4a)
plot((m4a), ask = FALSE)
plot(conditional_effects(m4a), ask = FALSE)
```
* Some evidence of zero-inflation (zi CI does not include 0).

#### Compare fit of models with and without zero inflation
```{r}
loo(m4, m4a)
```
* Zero inflated model somewhat better (elpd_diff/se_diff = 3.2).  

#### LOO probability integral transform
```{r}
loo_m4a <- loo(m4a, save_psis = TRUE)
```
```{r}
yrep <- posterior_predict(m4a)
ppc_loo_pit_overlay(
  y = d3a$bd_load, 
  yrep = yrep,
  lw = weights(loo_m4a$psis_object))
```
* Some improvement in uniformity (except at smallest values of x). 

#### Add expt_trt as shape predictor
```{r}
m4b <- brm(bf(bd_load ~ days * frog_group, shape ~ expt_trt),
          data = d3a,
          family = zero_inflated_negbinomial())
summary(m4b)
plot((m4b), ask = FALSE)
plot(conditional_effects(m4b), ask = FALSE)
```
* frog_group not an important predictor of shape (based on CI including 0). m4a is best. 

#### Plot of Bd loads over time 
```{r}
anno_text2 <- tibble(expt_trt = "control", bd_load = -0.81, lab = "text",
                    days = factor(56, levels = c("-7", "0", "12", "37", "56")))
```
```{r}
facet_names2 <- c('-7' = "Day -7",
                 '0' = "Day 0",
                 '12' = "Day 12",
                 '37' = "Day 37",
                 '56' = "Day 56")
```
```{r}
d %>%
  mutate(
    expt_trt = replace(expt_trt, is.na(expt_trt) & expt_or_wild == "wild", "wild")) %>% 
  ggplot(aes(x = expt_trt, y = log10(bd_load + 1), fill = expt_trt)) + 
    geom_boxplot(width = 0.9) +  
    scale_fill_manual(values = c("forestgreen", "deepskyblue2", "darkgray")) +
    labs(fill = "Frog group", 
         x = "Frog group",
         y = expression("Bd load"~(log[10]~("copies + 1")))) + 
    stat_n_text(size = 3) +
    geom_text(data = anno_text2, label = "n=0", size = 3) +
    facet_grid(~ days, labeller = as_labeller(facet_names2)) +
    theme_classic() +
    theme(axis.text = element_text(color = "black", angle = 90),
          axis.title.x = element_text(vjust = 0),
          legend.position = "none")
```
```{r}
ggsave(here::here("out", "figures", "dusy-2012-bdload.png"), width = 15, height = 10, units = "cm")
```

### ANALYSIS #5: Counts of recaptured animals in control, treated, and wild groups in 2012
#### Create dataset
```{r}
d4 <- d %>%
  left_join(jliv_days, by = "group_date") %>% 
  filter(group_date > "2012-07-19" & expt_or_wild == "expt") %>%
  mutate(
    released = case_when(
            expt_trt == "control" ~ 60,    # sample sizes from information recorded in field notebooks
            expt_trt == "treated" ~ 254))

d4_count <- d4 %>%
  group_by(days, expt_trt, released) %>%
  count(expt_trt) %>%
  ungroup() %>%
  complete(days, expt_trt, fill = list(n = 0, released = 60)) %>%
  mutate(freq = (n / released))
```
* Too few recaptures of control subadults (due to low apparent survival) to allow formal analysis

#### Plot showing frog recaptures over time in 2012
```{r}
d4_count %>% 
  ggplot(aes(x = days, y = freq, fill = expt_trt)) +
    geom_line() +
    geom_point(shape = 21, size = 4) + 
    scale_fill_manual(values = c("forestgreen", "deepskyblue2")) +
    labs(
      fill = "Treatment group",
      x = expression("Days since"~italic("J. lividum")~"exposure"),
      y = "Percent of frogs recaptured") +
    theme_classic() +
    theme(axis.text = element_text(color = "black"),
          legend.position = c(0.80, 0.85),
          legend.background = element_rect(size = 0.3, color = "black"))
```

```{r}
ggsave(here::here("out", "figures", "dusy-2012-recapture.png"), width = 10, height = 10, units = "cm")
```
