---
title: "Dusy 2010 zoospore pool data"
author: "Roland Knapp"
output: 
editor_options: 
  chunk_output_type: console
---
## Dataset provided by Andrea Jani

## ANALYSIS #1: Effect of 2010 Dusy Basin itraconazole treatment on Bd zoospore pool
## Load packages
```{r setup}
library(brms)
library(loo)
library(bayesplot)
library(tidyverse)
```

### Read in data. To minimize effect of false negatives, removed all filters with Bd load = 0
```{r}
d1 <- read_csv(here::here("data", "dusy-2010-zsppool.csv"))
```
```{r}
d1 <- d1 %>% 
  filter(bd_load > 0) %>% 
  mutate(pre_post = fct_relevel(pre_post, "pre", "post"),
         tmt = as.factor(tmt)) %>% 
  select(-group)
```

### Data exploration
```{r}
d1 %>% 
  ggplot(aes(x = tmt, y = log10(bd_load), fill = tmt)) +
  geom_boxplot() +
  facet_grid(.~ pre_post)
```

### Distribution of bd_load 
```{r}
d1 %>%
  ggplot(aes(x = bd_load)) +
    geom_histogram(bins = 100)
```
```{r}
d1 %>% filter(sample_id != "AJF132") %>%
  ggplot(aes(x = bd_load)) +
    geom_histogram(bins = 50)
```
```{r}
d1 %>% filter(sample_id != "AJF132") %>%
  ggplot(aes(x = log10(bd_load))) +
    geom_histogram(bins = 50)
```
* filter AJF132 has loads an order of magnitude higher than other filters. 
* Data are left-skewed. 

### Fit negative binomial model
```{r}
d1a <- d1 %>% 
  mutate(bd_load = as.integer(bd_load))
```
```{r}
m1 <- brm(bd_load ~ pre_post * tmt + (1 | sample_id),
          data = d1a,
          family = negbinomial(),
          save_all_pars = TRUE) # for loo below - deprecated but works (new format "save_pars(all = TRUE)" produces error)
summary(m1)
plot(m1, ask = FALSE)
plot(conditional_effects(m1), ask = FALSE)
```
* Effects of pre_post, trt, and interaction are all unimportant.
* Variation in load is higher for "post" vs "pre", and "treated" vs "control" samples.

#### Compute fit diagnostics
```{r}
loo1 <- loo(m1, save_psis = TRUE)  # 17 observations with pareto_k > 0.7
```
```{r}
loo1 <- loo(m1, save_psis = TRUE, moment_match = TRUE) # 2 observations with pareto_k > 0
```
```{r}
yrep <- posterior_predict(m1)
ppc_loo_pit_overlay(
  y = d1a$bd_load, 
  yrep = yrep,
  lw = weights(loo1$psis_object))
```
* LOO-PIT has non-uniform distribution.

### Add predictors of shape
```{r}
m1a <- brm(bf(bd_load ~ pre_post * tmt + (1 | sample_id), shape ~ pre_post + tmt),
          data = d1a,
          family = negbinomial())
summary(m1a)
plot(m1a, ask = FALSE)
plot(conditional_effects(m1a), ask = FALSE)
```
* Predictors of shape are both unimportant.
* No additional analyses possible. 
* Small sample sizes are likely a factor in poor fit. 

## CREATE PLOT 
```{r}
library(EnvStats)
```
```{r}
prepost_names <- c('pre' = "Before treatment",
                   'post' = "After treatment")
```
```{r}
d1 %>% 
  ggplot(aes(x = tmt, y = log10(bd_load), fill = tmt)) +
  geom_dotplot(method = "histodot", binaxis = "y", stackdir = "center", stackratio = 1, dotsize = 0.8, stroke = 0.5) + 
   facet_grid(~ pre_post, labeller = as_labeller(prepost_names)) +
   labs(
     x = "Frog group", 
     y = expression("Bd load"~(log[10]("copies")))) +
   scale_fill_manual(values = c("forestgreen", "deepskyblue2")) +
   scale_x_discrete(labels = c("Control", "Treated")) +
   stat_n_text(size = 3) + 
   stat_summary(fun = median, geom = "point", shape = 18, size = 2.5, show.legend = FALSE) +
   theme_classic() +
   theme(axis.text = element_text(color = "black"),
         axis.title.x = element_text(vjust = 0.5),
         legend.position = "none")
```
```{r}
ggsave(here::here("out", "figures", "dusy-zsppool.png"), width = 10, height = 10, units = "cm")
```



