---
title: "LeConte treatment"
author: "Roland Knapp"
date: "12/20/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(brms)
library(tidyverse)
library(loo)
library(bayesplot)
```

# Create dataset for analysis
## Read in data
```{r}
d <- read_csv(here::here("data", "leconte-20152018-captures.csv"))
```

## Select 2015 data (collected during treatment period)
```{r}
d <- d %>%
  filter(visit_date < "2015-12-31") %>%
  drop_na(bd_load) %>% 
  mutate(bd_load = as.integer(bd_load)) %>%
  rename(group = category) %>% 
  select(location, group, trt_period, visit_date, trt_died, pit_tag_ref, bd_load) %>% as_tibble()
```

# ANALYSIS #1:  Bd load between treated and control groups before treatment
## Create dataset, check distribution of response variable
```{r}
d1 <- d %>% filter(trt_period == "pretreat" & trt_died == FALSE) %>%
  mutate(across(c("location", "group"), as.factor))

d1 %>%
  ggplot(aes(x = bd_load)) +
    geom_histogram(bins = 200)
```
* Load data are strongly left-skewed, may be zero-inflated

## Negative binomial model
```{r}
m1 <- brm(bd_load ~ location * group,
          data = d1,
          family = negbinomial)
summary(m1)
plot((m1), ask = FALSE)
plot(conditional_effects(m1), ask = FALSE)  
```
* Important effect of group, effect of location is unimportant, weak effect of interaction

#### LOO probability integral transform
```{r}
loo1 <- loo(m1, save_psis = TRUE)
yrep <- posterior_predict(m1)
ppc_loo_pit_overlay(
  y = d1$bd_load, 
  yrep = yrep,
  lw = weights(loo1$psis_object))
```
* Fit of m1 is quite good, uniform distribution

## Zero-inflated negative binomial
```{r}
m1a <- brm(bd_load ~ location * group,
          data = d1,
          family = zero_inflated_negbinomial())
summary(m1a)
plot((m1a), ask = FALSE)
plot(conditional_effects(m1a), ask = FALSE)  
```
* No evidence of zero inflation (zi-estimate = 0), use negative binomial model

# ANALYSIS #2: Bd load of treated group before/after treatment
## Create dataset, check distribution of response variable
```{r}
d2 <- d %>% filter(group == "treated" & (trt_died == FALSE | is.na(trt_died))) %>%
  mutate(location = as.factor(location),
         trt_period = fct_relevel(trt_period, rev))
```

### Distribution of Bd load data
```{r}
d2 %>%
  ggplot(aes(x = bd_load)) +
    geom_histogram(bins = 200)
```
* Load data are strongly left-skewed, may be zero-inflated

## Negative binomial model
```{r}
m2 <- brm(bd_load ~ location * trt_period,
          data = d2,
          family = negbinomial)
summary(m2)
plot((m2), ask = FALSE)
plot(conditional_effects(m2), ask = FALSE)  
```
* Important effects of location (higher loads in upper), trt_period (lower after treatment), interaction (loads start higher & end higher in upper)

#### LOO probability integral transform
```{r}
loo2 <- loo(m2, save_psis = TRUE)
yrep <- posterior_predict(m2)
ppc_loo_pit_overlay(
  y = d2$bd_load, 
  yrep = yrep,
  lw = weights(loo2$psis_object))
```
* Fit of m2 is quite good, uniform distribution

## Zero-inflated negative binomial
```{r}
m2a <- brm(bd_load ~ location * trt_period,
          data = d2,
          family = zero_inflated_negbinomial())
summary(m2a)
plot((m2a), ask = FALSE)
plot(conditional_effects(m2a), ask = FALSE)  
```
* No evidence of zero inflation (zi-estimate = 0.03), use negative binomial model

# ANALYSIS #3: Loads of frogs that lived vs died during treatment
```{r}
d3 <- d %>% filter(trt_period == "pretreat" & group == "treated") %>% 
    mutate(across(c("location", "trt_died"), as.factor),
           bd_load = as.integer(bd_load)) %>% as_tibble()
```

### Distribution of Bd load data
```{r}
d3 %>% ggplot(aes(x = bd_load)) +
          geom_histogram(bins = 50)
```
* Strong negative skew, zero-inflation isn't extreme

### Fit negative binomial model
```{r}
m3 <- brm(bd_load ~ trt_died * location,
          data = d3,
          family = negbinomial())
summary(m3)
plot(m3)
plot(conditional_effects(m3), ask = FALSE)
```

### Fit negative binomial model with shape predictors
```{r}
m3b <- brm(bf(bd_load ~ trt_died * location, 
             shape ~ trt_died + location),
          data = d3,
          family = negbinomial())
summary(m3b)
plot((m3b), ask = FALSE)
plot(conditional_effects(m3b), ask = FALSE)
```
* Effects of predictors are unchanged.

#### Compare fit of models with and wo shape predictors
```{r}
loo3 <- loo(m3, save_psis = TRUE)
loo3b <- loo(m3b, save_psis = TRUE)

loo(m3, m3b)
```
* m3b has a better fit (diff ratio = 3.1), but see LOO-PIT results. Use simpler model, m3.

#### LOO probability integral transform
```{r}
# for m3b
yrep <- posterior_predict(m3b)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3b$psis_object))

# for m3
yrep <- posterior_predict(m3)
ppc_loo_pit_overlay(
  y = d3$bd_load, 
  yrep = yrep,
  lw = weights(loo3$psis_object))
```
* LOO-PIT shows better fit of m3 - more uniform

### Fit zero-inflated negative binomial model
```{r}
m3a <- brm(bd_load ~ trt_died * location,
          data = d3,
          family = zero_inflated_negbinomial())
summary(m3a)
plot((m3a), ask = FALSE)
plot(conditional_effects(m3a), ask = FALSE)
```
* No evidence of zero-inflation

# CREATE PLOT
## Load packages needed for plotting
```{r}
library(patchwork)
library(EnvStats)
```

## Plot of Bd loads before and at end of treatment, by location
```{r}
p1 <- d %>% 
  filter(location == "lower" & (trt_died == FALSE | is.na(trt_died))) %>% 
  mutate(trt_period = fct_relevel(trt_period, rev)) %>% 
  ggplot(aes(x = trt_period, y = log10(bd_load + 1), fill = group)) +
  geom_boxplot(width = 0.9, position = position_dodge2(preserve = "single")) + # keep unmatched boxplot same size as others
    labs(title = "(a)",
      x = "Treatment period",
      y = expression("Bd load"~(log[10]~("copies + 1")))) +
    scale_x_discrete(labels = c("Before","End")) + # rename x axis labels
    scale_fill_manual(values = c("forestgreen", "deepskyblue2")) + 
    ggtitle("(a)") +
    annotate("text", x = c("pretreat", "endtreat"), y = -0.5, label = c("n=102  n=285", "n=96"), size = 3) +
    theme_classic() +
    theme(legend.position = "none", 
          axis.text = element_text(color = "black"),
          axis.title.x = element_text(vjust = -0.7))

p2 <- d %>% 
  filter(location == "upper" & (trt_died == FALSE | is.na(trt_died))) %>% 
  mutate(trt_period = fct_relevel(trt_period, rev)) %>% 
  ggplot(aes(x = trt_period, y = log10(bd_load + 1), fill = group)) +
  geom_boxplot(width = 0.9, position = position_dodge2(preserve = "single")) + # keep unmatched boxplot same size as others
    labs(title = "(b)", 
      x = "Treatment period",
      y = expression("Bd load"~(log[10]~("copies + 1"))), 
      fill = "Treatment\ngroup") +
    scale_x_discrete(labels = c("Before","End")) + 
    scale_fill_manual(values = c("forestgreen", "deepskyblue2")) + 
    annotate("text", x = c("pretreat", "endtreat"), y = -0.5, label = c("n=74  n=125", "n=50"), size = 3) +
    theme_classic() + 
    theme(axis.text = element_text(color = "black"),
          axis.title.x = element_text(vjust = -0.7),
          legend.position = c(0.3, 0.35),
          legend.background = element_rect(size = 0.3, color = "black"))

fig4 <- p1 + p2
fig4
```
```{r}
ggsave(here::here("out", "figures", "leconte-2015-beforeafter-treat.png"), width = 15, height = 10, units = "cm")
```



